---
title: 'Reproducible Research : Peer Assessment 1'
author: "YOU"
date: "July 18, 2019"
output: html_document
---
library(knitr)
library(dplyr)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



##Loading and preprocessing the data

 #Loading data

```{r}
data_row <- read.csv("activity.csv", na.strings = c("NA"))
data <- data_row[with(data_row, {!(is.na(steps))}), ]
head(data,20)

```


##what is mean total number of steps taken per day?
```{r}
by_day <- group_by(data, date)
steps_by_day <- summarise(by_day, total = sum(steps))
steps_by_day
```

 #Ploting the Histogram
```{r}
library(ggplot2)
qplot(steps_by_day$total, main = "Total number of steps taken each day", xlab = "Total Steps", ylab = "Frequency", col = "blue")
```

 #calculating the mean and median of the total number of steps taken per day
```{r}
summary(steps_by_day)
```

Mean of total number of steps taken per day is 10766 and median is 20765.


##What is the average daily activity pattern?
 #preprocessing data for the plot
```{r}
steps_by_interval <- aggregate(steps~interval, data, mean)
```

 #create a time series plot
```{r}
plot(steps_by_interval$interval, steps_by_interval$steps, type ="l", main = "Average number of steps overl all days", xlab = "Interval", ylab = " Average number of steps")
```

 #finding the max number of steps
```{r}
max_steps_row <- which.max(steps_by_interval$steps)
steps_by_interval[max_steps_row,]
```

The interval 835 has the maximum average value of steps(206.1698)
 
##Inputing missing values

 #Total no. of rows with NA's
```{r}
sum(is.na(data_row))
```
Total number of rows with NA's is 2304
 
 #I picked the strategy of replacing NA's with the 5-minute interval
```{r}
data_imputed <- data_row
for(i in 1:nrow(data_imputed)){
  if(is.na(data_imputed$steps[i])){
  interval_value <- data_imputed$interval[i]
  steps_value <- steps_by_interval[
    steps_by_interval$interval == interval_value, ]
  data_imputed$steps[i] <- steps_value$steps
  }
}
```

 #I have created a new data set with out NA's. All NA's are replaces with mean of 5-minute interval.
```{r}
df_imputed_steps_by_day <- aggregate(steps~date, data_imputed, sum)
head(df_imputed_steps_by_day)
```

 #plotting the histogram
```{r}
qplot(df_imputed_steps_by_day$steps, main = "Total number of steps taken each day", xlab = "Total number of steps in a day", ylab ="Frequency", binwidth = 1000)
```

 #Calculating the mean and median values of imputed data
```{r}
mean(df_imputed_steps_by_day$steps)
```
 mean value is 10766.19
 
```{r}
median(df_imputed_steps_by_day$steps)
```
 median value is 10766.19
 
 #Calculatinf mean and median values without NA's
```{r}
mean(steps_by_day$total)
```
 mean value is 10766.19

```{r}
median(steps_by_day$total)
```
 median value is 10765
 
 Mean values stays the same but there is a slight difference in the median value.
 
 
##Are there differences in activity patterns between weekdays and weekends?
```{r}
data_imputed["type_of_day"] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day %in% c("Saturday", "Sunday")] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
```

 #convert type_of_day from character to factor
```{r}
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)
```

 #calculate avg steps by interval across all days
```{r}
df_imputed_steps_by_interval <- aggregate(steps~interval + type_of_day, data_imputed, mean)
```

 # Plotting
```{r}
qplot(interval, steps, data = df_imputed_steps_by_interval, type = "l", geom = c("line"), xlab = "Interval", ylab = "Number of steps", main = "")+
+ facet_wrap(~type_of_day, ncol = 1)
```

 
